---
title: "Simulation Modeling"
author: "Shannon Hagerty"
date: "8/12/2019"
output: html_document
---
## Simulation Modeling in R 

So today we're going to be building models to track stuff over time.  We're going to be building ordinary differential equations and then graphing the outputs. We'll use the package deSolve for this. 

```{r}
#install.packages('deSolve')
library(deSolve)
library(tidyverse)
```

```{r}
help(deSolve)
```

#### Tracking something over time

Our first model is going to be tracking a mortgage over time.  I think mortgages are confusing. When you first set it up you sometimes have the option of paying to lower your interest rate, alternatively you can add money to your downpayment, and you can even pay more than the minimum balance.I found it was easier to understand how this all worked out when I actually modeled it and saw the change in how fast the loan is paid off..

As an overview, when you have a mortgage each month you make a mortgage payment AND interest is accrued based on your interest rate.  How much of the total loan amount you still owe is a balance between your paying it off and the interest being charged.

First thing we do is set the time we want to model.  We're going to model over 30 years and we're looking at monthly changes so we'll model it over 360 months. 
```{r}
model_time <- seq(0, 360, by =1) 
```
seq() just creates a vector of values ranging from the first parameter to the second parameter incremented by whatever you say by is = to. 

Now we set the parameters of our model, the annual interest rate (divided over 12 months), and the monthly payment we make. 
```{r}
parameters <- c(interest_rate = (0.041/12), monthly_mortgage = 1095.89)
```
The state refers to the thing we are tracking.  In this case its the starting balance we owe on our mortgage, the principal. Let's also track the amount we've paid.  At the start of the mortgage this is 0. 
```{r}
state <- c(principal = 226800)
```

Next we use a function to build a model that describes how our principal balance changes over time. The with function 
```{r}
mortgage_model<-function(times,state,parameters) {
  with(as.list(c(state, parameters)), {
    dPrincipal =(interest_rate* principal) - monthly_mortgage
        list(c(dPrincipal))
  })
}
```

We run the model using the ode function and then save that output as dataframe called mortgage_model_data 
```{r}
mortgage_model_output<- ode(y = state, times = model_time, func = mortgage_model, parms = parameters)

mortgage_model_data<-data.frame(mortgage_model_output)
```
 


```{r}
ggplot(mortgage_model_data)+
  geom_line(aes(x=time, y=principal), size = 1.5, color='magenta')+
  theme_classic()+
  xlab('Months')+
  ylab('Mortgage')
```

# Try to make some changes. 
What if you pay an extra $100 a month toward your mortgage? (Below is the same code from the previous example. You will only need to change one piece of the code below then re run the chunk, you may want to rename variables so you don't overwrite the last example.)
```{r}
# Set Model time
model_time2 <- seq(0, 360, by =1) 

# Set model parameters
parameters2 <- c(interest_rate = (0.041/12), monthly_mortgage = 1095.89)

# Set model starting values
state2 <- c(principal = 226800)

# Set model equations
mortgage_model<-function(times,state,parameters) {
  with(as.list(c(state, parameters)), {
    dPrincipal =(interest_rate* principal) - monthly_mortgage
        list(c(dPrincipal))
  })
}

# Run model and generate output
mortgage_model_output2<- ode(y = state2, times = model_time2, func = mortgage_model, parms = parameters2)

mortgage_model_data2<-data.frame(mortgage_model_output2)
```
Now plot the data and see when does the line hit $ 0 balance? 
```{r}
#plot model output 
```

# Another model simulation
What if we decide to pay another $10,000 toward our loan right at the start, but leave the monthly payment the same.  How much does changing the initial amount owed alter the rate of total payment? Make the necessary adjustments and find out.  
```{r}
# Set Model time
model_time3 <- seq(0, 360, by =1) 

# Set model parameters
parameters3 <- c(interest_rate = (0.041/12), monthly_mortgage = 1095.89)

# Set model starting values
state3 <- c(principal = 226800)

# Set model equations
mortgage_model<-function(times,state,parameters) {
  with(as.list(c(state, parameters)), {
    dPrincipal =(interest_rate* principal) - monthly_mortgage
        list(c(dPrincipal))
  })
}

# Run model and generate output
mortgage_model_output3<- ode(y = state3, times = model_time3, func = mortgage_model, parms = parameters3)

mortgage_model_data3<-data.frame(mortgage_model_output3)
```
Okay now plot this output to see what the rate of payment looks like
```{r}

```
#### Wrangling Exercise
Let's generate a plot that shows all three scenarios at once.
for each mortgage_model_data we created, add a column called scenario and fill it with 'typical' (i.e mortgage_model_data), 'higher payment' (for mortgage_model_data2), 'increase downpayment' (for mortgage_model_data3). 
```{r}
#  

```
Next lets join all three data sets. We just want to paste them together to make one really long data set. Use bind_rows(), inside the parentheses call the three dataframes separating them by commas. Save to a variable name. 
```{r}

```
Take your new combined dataframe and plot all three lines, you can do this by setting color=scenario inside the aes() in geom_line()
```{r}

```

## Tracking more than one thing over time 

```{r}
# Set Model time
LV_time <- seq(0, 100, by =1) 

# Set model parameters
LV_parameters <- c(
  prey_growth = 2,
  attack_rate = .5,
  predator_death = .2,
  predator_efficiency = .4)

# Set model starting values
LV_state <- c(prey = 10, 
              predator =10)

# Set model equations
LV_model<-function(times,state,parameters) {
  with(as.list(c(state, parameters)), {
    dPrey = prey_growth*prey - attack_rate*prey*predator
    dPredator = predator_efficiency*prey*predator - predator_death*predator
        list(c(dPrey, dPredator))
  })
}

# Run model and generate output
LV_output<- ode(y = LV_state, times = LV_time, func = LV_model, parms = LV_parameters)

LV_data<-data.frame(LV_output)
```
In order to plot the data we'll need to restructure the dataframe so that there is a column with the count of predator or prey and another column indicating the population (i.e. predator or prey). That way we can use the population column to make the lines plot with different colors in ggplot. 
```{r}
#LV_data_gather <- gather(#use documentation to identify how to gather the data)
```
Now plot it
```{r}
#plot!
```

