---
title: "Statistical Models/Tests Part2"
author: "Margarita Corral"
date: "7/22/2019"
output: html_document
---
Welcome to our second session on statistical analysis in R!  

### Today we will cover:

- Linear regression
- Checking linear regression assumptions
- Logistic regression

### Set up
To get started let's load or install the library's we'll be using. If this is your first time using one of the packages uncomment and run the appropriate install.package('package')
```{r}
#install.packages('tidyverse')
library(tidyverse)

install.packages("GGally")
library(GGally)

install.packages("car")
library(car)

install.packages("lmtest")
library(lmtest)

install.packages('texreg')
library(texreg)


```

We're going to be using the "states" data set. 




```{r}
States <-read_csv("states.csv")
```


## Let's explore the data

```{r}
View(States)
```

```{r}
summary(States)
```


Before fitting a model to predict price, let's take a look at our dependent variable 
density plot
```{r}
States %>% 
  ggplot(aes(cons_hr)) +
  stat_density() + 
  theme_bw()
```

correlation matrix
```{r}
ggcorr(States)
```
correlation matrix
```{r}
ggcorr(States[,3:14], label=TRUE)
```

correlation
```{r}
cor(States$cons_hr, States$college)

```


Univariate model
```{r}
model1<-lm(cons_hr ~ college, data=States)

summary(model1)

```
We can use the screenreg function from texreg to get a nice table

```{r}
screenreg(model1)
```

```{r}
screenreg(model1, 
          custom.model.names = "Model 1",  
          custom.coef.names = c("Constant", "% of college graduates"))
```
We can export the table as .doc
```{r}
htmlreg(list(model1), file = "model_1.doc", 
    custom.model.names = "Model 1",  
    custom.coef.names = c("Constant", "% of college graduates"),
    inline.css = FALSE, doctype = TRUE, html.tag = TRUE, 
    head.tag = TRUE, body.tag = TRUE)
```



Multivariate model
```{r}
model2<-lm(cons_hr ~ college+christad, data=States)
```

```{r}
screenreg(model2)
```

Linear Regression Assumptions

1. Residuals are normally distributed. The plot()funciton will give us 4 diagnostic plots
```{r}
res<-residuals(model2)
res <- as.data.frame(res)
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5,bins=15)
plot(model2)

```
```{r}
shapiro.test(model2$residuals)

```



2. Homoscedasticity. we can check the Residuals vs fitted plot or use the Breusch-Pagan test (null hypothesis: heteroskedasticity is not present)
```{r}
bptest(model2)
```

3.Multicollinearity. We can use vif() to calculate the variance inflation factor. A value larger than 2.5 may be a cause for concern
```{r}
vif(model2)
```

Model with categorical variables
```{r}
model3<-lm(cons_hr ~ college+christad+region, data=States)
summary(model3)
```

Changing the reference goup
```{r}
States$region<-as.factor(States$region)
States$region<-relevel(States$region, ref="South")
model4<-lm(cons_hr ~ college+christad+region, data=States)
summary(model4)
```
Modeling interactions
```{r}
model5<- lm(cons_hr ~ college*region, data=States)
summary(model5)
```


Logistic regression
```{r}
str(States$gb_win04)
States$gb_win04<-as.factor(States$gb_win04)
```

```{r}
model6<-glm(gb_win04~prcapinc, data=States, family=binomial)
summary(model6)
```







[This week's TidyTuesday:R for Data Science Online Learning Community Stats]
```{r}

r4ds_members <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-16/r4ds_members.csv")

```

